---
title: "Final_Project"
author: "Deekshith Reddy Bhoomireddy, Milton La"
date: "2024-12-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Front Matter - Clean Environment, Load Libraries, User Defined Functions

```{r}
rm(list = ls())
#Add libraries as needed
library(tidyverse)
library(readr)
library(stringr)
library(caret) # For data splitting and model evaluation


player1 <- read_csv("CODGames_p1_380.csv", show_col_types = FALSE)
player2 <- read_csv("CODGames_p2_380.csv", show_col_types = FALSE)
maps <- read_csv("CODMaps.csv", show_col_types = FALSE)
modes <- read_csv("CODGameModes.csv", show_col_types = FALSE)
```

## Task 1
#### Data Cleaning

```{r}
#Combining player1 and player2 data
# Step 1: Add a unique key for each dataset
new_p1 <- player1 |>
  mutate(PlayerID = "Player1")

new_p2 <- player2 |>
  mutate(PlayerID = "Player2")

# Step 2: Perform a full join based on shared columns (excluding PlayerID)
combined_data <- merge(new_p1, new_p2, all = TRUE, by = intersect(names(new_p1), names(new_p2)))
```


```{r}
#Combining combined data with maps data

#Corrections in Choice column in Combined data
combined_data <- combined_data |>
  mutate(Choice = str_replace_all(Choice, "Deisel", "Diesel"),
         Choice = str_replace_all(Choice, "Apocolypse", "Apocalypse"),
         Choice = str_replace_all(Choice, "APocalypse", "Apocalypse"),
         Choice = str_replace_all(Choice, "^Collateral$", "Collateral Strike"),
         Choice = str_replace_all(Choice, "Collaterel Strike", "Collateral Strike"),
         Choice = str_replace_all(Choice, "Riad", "Raid"),
         Choice = str_replace_all(Choice, "Drive-in", "Drive-In"),
         Choice = str_replace_all(Choice, "Nuketown '84 Halloween", "Nuketown '84"))

# Step 1: Rename the Date column in maps_data to MapDate
maps <- 
  maps |>
  rename(MapDate = Date)

# Step 2: Perform a left join to add map details
maps_combined_data <- combined_data |>
  left_join(maps, by = c("Choice" = "Name"))
```



#### Data Visualization
```{r}

# Step 1: Count occurrences as candidates
map_count <- 
  maps_combined_data |>
  pivot_longer(cols = c(Map1, Map2), names_to = "Chosen", values_to = "Map") |>
  filter(!is.na(Map) & Map != "") |> 
  group_by(Map) |>
  summarize(MapCount = n())

# Step 3: Count occurrences as winners
map_won <- 
  maps_combined_data |>
  group_by(Choice) |>
  summarize(WinCount = n()) |>
  rename(Map = Choice)

# Step 4: Merge candidate and win counts
map_statistics <- 
  map_count |>
  left_join(map_won, by = "Map") |>
  filter(!is.na(WinCount)) |>  # Remove rows with NA WinCount
  mutate(WinProportion = WinCount / MapCount) |>
  arrange(desc(WinProportion))

# Step 5: Visualization of results
ggplot(map_statistics, aes(x = reorder(Map, WinProportion), y = WinProportion)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Winning Probability of Maps in Map Vote",
    x = "Map",
    y = "Winning Proportion"
  )

# Display map statistics
head(map_statistics)
```
The analysis of map vote winning proportions reveals that Drive-In and Nuketown '84 stand out with proportions exceeding 1. This outstanding proportions indicate that they were often chosen either due to player preference, game mechanics (such as tie breaker mechanism), or lobby returns. Standoff achieved a perfect proportion of 1, indicating it was always chosen when available. Other all maps such as Crossroads Strike, Diesel and Raid displayed high popularity with proportions above 0.5, showing that they were often strong contenders but occasionally lost to other options. Other all maps with proportions below 0.5 are often selected when the players do not have desirable options or a worst other option showing that they often lost to other options.




## Task 2
#### Data Cleaning
```{r}
# Define a reusable correction function
correction_values <- c(
  "Deisel" = "Diesel",
  "Apocolypse" = "Apocalypse",
  "APocalypse" = "Apocalypse",
  "Collateral" = "Collateral Strike",
  "Collaterel Strike" = "Collateral Strike",
  "Riad" = "Raid",
  "Drive-in" = "Drive-In",
  "Nuketown '84 Halloween" = "Nuketown '84"
)

apply_corrections <- function(dataset, column) {
  dataset %>%
    mutate({{ column }} := recode({{ column }}, !!!correction_values))
}

# Apply corrections to datasets
data_player1 <- apply_corrections(player1, Choice)
data_player2 <- apply_corrections(player2, Choice)
```


```{r}
# Combine player1 and player2 data
data_p1 <- data_player1 %>% mutate(PlayerLabel = "Player1")
data_p2 <- data_player2 %>% mutate(PlayerLabel = "Player2")

merged_data <- full_join(data_p1, data_p2, by = intersect(names(data_p1), names(data_p2)))

# Merge merged_data with map details
combined_maps_data <- merged_data %>%
  left_join(maps, by = c("Choice" = "Name"))
```


#### Data Visualization

```{r}
# Count occurrences as candidates
map_candidates <- combined_maps_data %>%
  pivot_longer(cols = c(Map1, Map2), names_to = "Chosen", values_to = "Map") %>%
  filter(!is.na(Map) & Map != "") %>%
  group_by(Map) %>%
  summarize(MapCount = n())

# Count occurrences as winners
map_wins <- combined_maps_data %>%
  group_by(Choice) %>%
  summarize(WinCount = n()) %>%
  rename(Map = Choice)

# Merge candidate and win counts
map_stats <- map_candidates %>%
  left_join(map_wins, by = "Map") %>%
  filter(!is.na(WinCount)) %>%
  mutate(WinProportion = WinCount / MapCount) %>%
  arrange(desc(WinProportion))

# Visualization of results
ggplot(map_stats, aes(x = reorder(Map, WinProportion), y = WinProportion)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Winning Probability of Maps in Map Vote",
    x = "Map",
    y = "Winning Proportion"
  )

# Display sorted map statistics
print(map_stats)
```

#### Write the comparision of task 1 and task 2 and the conclusion




## Task 3

#### Data Cleaning
```{r}
# Replace "None" with NA in TimeLimit and clean GameType
modes <- modes %>% mutate(TimeLimit = na_if(TimeLimit, "None"))

maps_combined_data <- maps_combined_data %>%
  mutate(GameType = recode(GameType,
                           "HC - TDM" = "TDM",
                           "HC - Hardpoint" = "Hardpoint",
                           "HC - Kill Confirmed" = "Kill Confirmed",
                           "HC - Domination" = "Domination"))

final_data <- maps_combined_data %>%
  left_join(modes, by = c("GameType" = "Mode"))
```


```{r}
# Filter out rows with missing or non-finite values
cleaned_data <- final_data %>%
  filter(!is.na(Score) & !is.na(TotalXP) & is.finite(Score) & is.finite(TotalXP))
```

```{r}
summary_stats <- cleaned_data %>%
  group_by(GameType) %>%
  summarize(
    mean_Score = mean(Score, na.rm = TRUE),
    sd_Score = sd(Score, na.rm = TRUE),
    mean_TotalXP = mean(TotalXP, na.rm = TRUE),
    sd_TotalXP = sd(TotalXP, na.rm = TRUE),
    n = n()
  )

print(summary_stats)
```

```{r}
ggplot(cleaned_data, aes(x = Score, y = TotalXP, color = GameType)) +
  geom_point(alpha = 0.5) +  # Scatter points
  geom_smooth(method = "loess", se = FALSE) +  # Add a LOESS curve
  facet_wrap(~ GameType) +  # Separate by GameType
  labs(title = "Checking Linearity Between Score and TotalXP",
       x = "Score",
       y = "TotalXP")
```

```{r}
# Linear Regression Model
model <- lm(TotalXP ~ Score * GameType, data = cleaned_data)
summary(model)
```
```{r}
# Residual diagnostics
par(mfrow = c(2, 2))
plot(model)
```






## Task 4

