---
title: "Mini Project 1"
author: "Milton La, Deekshith Reddy"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter
```{r}
library(readxl)
library(tidyverse)
library(lubridate)
data <- read_xlsx("CODGames2_mp.xlsx")
```
## Exploring the Dataset

#### Selected Variables:
Categorial Variable: PrimaryWeapon, GameType
Quantitative Variable: Eliminations, TotalXP, Score

```{r}
# Summary statistics for a categorial variable: "PrimaryWeapon"
PrimaryWeapon <- data$PrimaryWeapon
table(PrimaryWeapon)
prop.table(table(PrimaryWeapon))

# Data visualization for "PrimaryWeapon"
data |>
  ggplot(aes(x = PrimaryWeapon)) +
  geom_bar(fill = "darkred", color = "black") + 
  labs(x = "Primary Weapon", y = "Count")

# Summary statistics for a categorial variable: "GameType"
GameType <- data$GameType
table(GameType)
prop.table(table(GameType))

# Data visualization for "GameType"
data |>
  ggplot(aes(x = GameType)) +
  geom_bar(fill = "darkgreen", color = "black") + 
  labs(x = "Game Type", y = "Count")

# Summary statistics for a quantitative variable: "Eliminations"
Eliminations <- data$Eliminations
summary(Eliminations)
mean(Eliminations, na.rm = TRUE)
median(Eliminations, na.rm = TRUE)
max(Eliminations, na.rm = TRUE)
sd(Eliminations, na.rm = TRUE)

# Data visualization for "Eliminations"
data |>
  ggplot(aes(x = Eliminations)) +
  geom_boxplot(fill = "beige", color = "black") + 
  labs(x = "Number of Eliminations", y = "Count")

# Summary statistics for a quantitative variable: "TotalXP"
TotalXP <- data$TotalXP
summary(TotalXP)
mean(TotalXP, na.rm = TRUE)
median(TotalXP, na.rm = TRUE)
max(TotalXP, na.rm = TRUE)
sd(TotalXP, na.rm = TRUE)

# Data visualization for "TotalXP"
data |>
  ggplot(aes(x = TotalXP)) +
  geom_histogram(fill = "lightblue", color = "black", bins = 30) + 
  labs(x = "Total XP", y = "Frequency")

# Summary statistics for a quantitative variable: "Score"
Score <- data$Score
summary(Score)
mean(Score, na.rm = TRUE)
median(Score, na.rm = TRUE)
max(Score, na.rm = TRUE)
sd(Score, na.rm = TRUE)

# Data visualization for "Score"
data |>
  ggplot(aes(x = Score)) +
  geom_histogram(fill = "orange", color = "black", bins = 30) + 
  labs(x = "Score", y = "Frequency")
```


## Potential questions to ask:
-> How were the data for each match collected?
-> Are there any problem with the data (like missing values and any problems) which should be accounted for?
-> Were all the matches, from which data was collected, part of the same session.


## Part 1 - Is the player’s performance, as quantified by the amount of experience points gained (TotalXP variable) changing over time?
```{r}
data$Date <- as.Date(data$Date)
## Adding month to the data
data <- mutate(data, Month = month(Date, label = TRUE, abbr = FALSE))

## Side-by-side boxplot using "TotalXP"
ggplot(data, aes(x = Month, y = TotalXP)) +
  geom_boxplot(fill = "beige", color = "black") +
  labs(x = "Months", y = "TotalXP")
  
```
Yes, the player's performance is changing over time but the change is variables as the median performance decreases from June to July but increases from July to August.


## Part 2: Is the there a relationship between the damage and the number of eliminations?

```{r}
ggplot(data, aes(x = Damage, y = Eliminations)) +
  geom_point(color = "darkred") +
  geom_smooth(method = "lm", color = "gold", se = FALSE, formula = y~x)
  labs(x = "Damage", y = "Number of Eliminations")
```



